Software Citation Principles

### 2016-08-21 sc-principles/rebuttal-2/rebuttal.tex

We have updated the manuscript to better match this position.

\begin{itemize}

\item \emph{``commit references are not guaranteed to be permanent'':
identifiers consisting of strong hash values *are* guaranteed to be permanent
in the sense that they can forever only represent what they were meant to
represent, i.e. the bits and bytes on which the hash was calculated (which, in
the case of Git, is the entire version history).}

\textbf{Our response:}
It is possible to overwrite\slash clobber a particular commit (e.g, using a
git force-push), or remove a repository entirely, and so access to that hash
may not be persistent.

@@

Regarding the peristence of GitHub vs.\ the DOI system, DOIs are an ISO standard
governed by a nonprofit foundation with a specific goal of persistence. Beyond
that, there are robust disaster recovery solutions in place for archivers and
publishers---the persistence of scholarly objects (and the DOI system) is
more certain than that of GitHub.

Furthermore, one of the goals behind creating the software citation principles
is increasing the prominence of software as an equal research product to
publications and data, which both use DOIs as a standard identifier and locator.
Thus, we feel it is important (in combination with our other stated reasons) to
recommend a similar standard for software; while other solutions (such as URL +
hash) certainly exist, suggesting a different---if equivalently suitable---standard
would serve to propagate the idea that software is a lesser research output.

\item \emph{``A particular version number/commit reference may not represent a
``preferred'' point at which to cite the software from the perspective of the
package authors.'': Hashes certainly don't prevent you from defining preferred
versions for future reference. But they come with the *possibility* to identify
all versions, which I would in fact consider an important feature. If I want to
make my research perfectly reproducible, I should be able to cite the exact
version that I used, whether that happens to be a preferred version or not.}

\textbf{Our response:}
As stated in our general response above and reflected in our changes to the
manuscript, we agree hashes (combined with a
repository URL) could be used when citing a version of software that does not
have an associated DOI.
We do expect the follow-on implementation working group to work further on these solutions.

@@

\item \emph{``When we talk about a ``Unique identifier'' in the document, we
are referring to unique, persistent, and machine-actionable identifiers such as
a DOI, ARK, or PURL; we do not believe a repository address with a version
number is sufficient.'' [from the rebuttal letter]: Identifiers that include
hashes are in important ways *more* unique, *more* persistent, and *more*
machine-actionable than DOIs or PURLs. Strong hashing algorithms guarantee
uniqueness (no other input can be constructed for the same hash), are in
practice probably more persistent (for example, there exist independent
complete copies of all GitHub content, so even if github.com closes for good,
the cited software could still be found), and can be automatically verified
against their supposed content. (The above holds for *strong* hash functions,
and Git's SHA-1 is no longer considered strong - even though not a single
collision has been found until now - but this doesn't invalidate the general
points.)}

\textbf{Our response:}
The working group focused on DOIs as the primary unique identifier\slash
locator to satisfy the principles of unique identification, persistence, and
accessibility. While strong hashing algorithms (which Git does not use) are
unique, they cannot currently be used with GitHub repo URLs (e.g.). In addition,
we prefer a persistent solution that doesn't depend on outside solutions if
(e.g.) GitHub goes away---the DOI system ensures persistence regardless of the
original ``owner''\slash naming authority.

### 2016-08-21 sc-principles/software-citation-principles.tex
changing software vs data discussion to point to github doc

The process of creating the software citation principles began by adapting
the FORCE11 Data Citation Principles~\cite{data-citation-principles}. These
were then modified based on discussions of the FORCE11 Software Citation
Working Group (see Appendix~\ref{app:wg_members} for members), information from
the use cases in \S\ref{sec:use_cases}, and the related work in
\S\ref{sec:related_work}.
We made the adaptations because software, while similar to data in terms of
not traditionally having been cited in publications, is also different than
data. Software can be considered a type of data, but the converse is not
generally true. In addition:

\begin{itemize}
\item Software can be used to express or explain concepts, unlike data.
\item Software is updated more frequently that papers or data\footnote{Datasets
can of course be frequently updated, but each update likely represents the
findings of a new or repeated study.}.
\item Software is executable, unlike papers or data.
\item Software suffers from a different type of bit rot than data---software
must be constantly maintained so that it continues to function as both the
hardware and software environments that it is used on changes, as developers
find and fix bugs, and as user requirements demand new features and capabilities.
\item Software is frequently built to use other software, leading to complex
dependencies, and these dependent software packages also frequently change.
\item Software is generally smaller than data, so a number of the storage and
preservation constraints on data don't apply to software.
\item Software teams can be large and multidisciplinary, and with varied roles
that overlap and sometimes make contributions ambiguous---some of which
can rise to a level equivalent to paper authorship.
\item The lifetime of software is generally not as long as that of data.
\end{itemize}

>>>>

The process of creating the software citation principles began by adapting
the FORCE11 Data Citation Principles~\cite{data-citation-principles}. These
were then modified based on discussions of the FORCE11 Software Citation
Working Group (see Appendix~\ref{app:wg_members} for members), information from
the use cases in \S\ref{sec:use_cases}, and the related work in
\S\ref{sec:related_work}.

We made the adaptations because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data. In the context of research (e.g., in science), the term ``data'' usually refers to electronic records of observations made in the course of a research study (``raw data'') or to information derived from such observations by some form of processing (``processed data''), as well as the output of simulation or modeling software (``simulated data''). 
Some confusion about the distinction between software and data comes in part from the much wider sense that the term ``data'' has in computing and information science, where it refers to anything that can be processed by a computer. In that sense, software is just a special kind of data.
Because of this, citing software is not the same as citing data.  A more general discussion about these distinctions in currently underway~\cite{software-vs-data}.
